---
title: 'Developing Data Products: Shiny Application with Reactive Input'
author: "Ve...n Sri..n"
date: "November 28-December 01, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
  transition: rotate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tableHTML)
library(xml2)
library(rvest)
library(dplyr)
library(RCurl)
library(lubridate)
library(curl)
```
##  Introduction 

<div class="blue2">
Developing Data product course focuses on three major areas (Shiny, Leaflets, ioslides presentation) for implementing useful data analysis functions.  

Shiny is a modern programming in R for communicating insights from data. Shiny is developed by RStudio and it is a web application framework. Leaflets: is a widely used open source JavaScript library used to build web mapping applications.  Leaflet allows developers to very easily display tiled web maps hosted on a public server, with optional tiled overlays. ioslides: Helps creating a slide show for the web based viewing.

I have developed an application which incorporates these learnings in a form of World Map with a pop-up presenting the Population and Life Expectancy projection for Year 2023 .

</div>
## Features Implemented

- Leaflet - Map with multiple functions including  focus, zoom level, geo location etc
- Shiny - ui and server with a drop down features (for selecting the country) and  check boxes (for selecting the Life Expectancy and Population details)
- Calculations - Read data from html pages (UN website for more update to date and accurate numbers), a simple formula to calculate the projections for Year 2013 and Life Expectancy determination
- Various R libraries are leveraged for standardization and easier data manipulations

## Data and Calculation

Data from multiple authentic websites are used for determining the population and life expectancy.  The United Nations publishes the data on periodic basis - the data is directly read from the website for most up to date information.   

```{r echo = TRUE}

population <- read_html("http://bit.ly/29Qqzm3") %>% html_nodes("table") %>% .[[2]] %>% html_table(trim = TRUE)
population <-population[, c(2,5,6)]
names(population) <- c("country", "Population 2016", "Population 2017")
head(population,6)
```
Assumptions:  Mortality rate will remain some what constant.  Fertility rate will also remain constant over next 5 years.  No significant people migration due to war or other natural calamities

## Learnings from the project

- Developing a shiny application with associated supporting documentation (ioslides). 
- Deploying the application on Rstudio's shiny server
- Using the javascripts with leaflet
- Developing a interactive application using (ui.R) and data manipulation (Server.R)
- Displaying reactive output displayed as a result of server calculations
- Designing simple prediction model
- Map rending based on various leaflet features

## Repository Details

- Application:  Shiny Application has been deployed in <https://vesr.shinyapps.io/DevDataProductsVersion5/>

- Github repository:  readme file, application, Server.R and ui.R are located in  <https://github.com/gitvesr/DevelopingDataProducts>

- Rpub:  Pitch for the application with clear instructions <http://rpubs.com/vesr/ShinyAppsPitch>


## References and Reading Materials

UN Population statistics published information  <https://esa.un.org/unpd/wpp/>

Quick Start Guide with Leaflet <http://leafletjs.com/examples/quick-start/>

Shiny App Development <https://shiny.rstudio.com/>
